### 2019-02-13 pres/pres.tex
pres. finish

\documentclass{beamer}
\usetheme{Malmoe}
\usecolortheme{seahorse}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{datavisualization}
\usetikzlibrary{datavisualization.formats.functions}

\usepackage{textpos}
\usepackage{csquotes}
\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{mathtools}
\usepackage{subdepth}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{svg}
\usepackage{xcolor,colortbl}
\usepackage{media9}
\usepackage{setspace}

\let\oldcite=\cite
\renewcommand{\cite}[1]{\textcolor[rgb]{.55,.55,.89}{\oldcite{#1}}}

\setbeamerfont{frametitle}{size=\small}
\addbibresource{../mscthesis.bib}
\addbibresource{pres.bib}
\captionsetup{justification=raggedright,singlelinecheck=false}

\makeatletter
\AtBeginDocument{%
  \let\l@algorithm\l@lstlisting
  \let\c@algorithm\c@lstlisting
  \let\thealgorithm\thelstlisting
  \renewcommand{\lstlistlistingname}{Algorithms and source code}%
}
\makeatother

\lstdefinelanguage{GLSL}
{
  sensitive=true,
  morekeywords=[1]{
    attribute, const, uniform, varying,
    layout, centroid, flat, smooth,
    noperspective, break, continue, do,
    for, while, switch, case, default, if,
    else, in, out, inout, float, int, void,
    bool, true, false, invariant, discard,
    return, mat2, mat3, mat4, mat2x2, mat2x3,
    mat2x4, mat3x2, mat3x3, mat3x4, mat4x2,
    mat4x3, mat4x4, vec2, vec3, vec4, ivec2,
    ivec3, ivec4, bvec2, bvec3, bvec4, uint,
    uvec2, uvec3, uvec4, lowp, mediump, highp,
    precision, sampler1D, sampler2D, sampler3D,
    samplerCube, sampler1DShadow,
    sampler2DShadow, samplerCubeShadow,
    sampler1DArray, sampler2DArray,
    sampler1DArrayShadow, sampler2DArrayShadow,
    isampler1D, isampler2D, isampler3D,
    isamplerCube, isampler1DArray,
    isampler2DArray, usampler1D, usampler2D,
    usampler3D, usamplerCube, usampler1DArray,
    usampler2DArray, sampler2DRect,
    sampler2DRectShadow, isampler2DRect,
    usampler2DRect, samplerBuffer,
    isamplerBuffer, usamplerBuffer, sampler2DMS,
    isampler2DMS, usampler2DMS,
    sampler2DMSArray, isampler2DMSArray,
  usampler2DMSArray, struct},
  morekeywords=[2]{
    radians,degrees,sin,cos,tan,asin,acos,atan,
    atan,sinh,cosh,tanh,asinh,acosh,atanh,pow,
    exp,log,exp2,log2,sqrt,inversesqrt,abs,sign,
    floor,trunc,round,roundEven,ceil,fract,mod,modf,
    min,max,clamp,mix,step,smoothstep,isnan,isinf,
    floatBitsToInt,floatBitsToUint,intBitsToFloat,
    uintBitsToFloat,length,distance,dot,cross,
    normalize,faceforward,reflect,refract,
    matrixCompMult,outerProduct,transpose,
    determinant,inverse,lessThan,lessThanEqual,
    greaterThan,greaterThanEqual,equal,notEqual,
    any,all,not,textureSize,texture,textureProj,
    textureLod,textureOffset,texelFetch,
    texelFetchOffset,textureProjOffset,
    textureLodOffset,textureProjLod,
    textureProjLodOffset,textureGrad,
    textureGradOffset,textureProjGrad,
    textureProjGradOffset,texture1D,texture1DProj,
    texture1DProjLod,texture2D,texture2DProj,
    texture2DLod,texture2DProjLod,texture3D,
    texture3DProj,texture3DLod,texture3DProjLod,
    textureCube,textureCubeLod,shadow1D,shadow2D,
    shadow1DProj,shadow2DProj,shadow1DLod,
    shadow2DLod,shadow1DProjLod,shadow2DProjLod,
    dFdx,dFdy,fwidth,noise1,noise2,noise3,noise4,
  EmitVertex,EndPrimitive},
  morekeywords=[3]{
    gl_VertexID,gl_InstanceID,gl_Position,
    gl_PointSize,gl_ClipDistance,gl_PerVertex,
    gl_Layer,gl_ClipVertex,gl_FragCoord,
    gl_FrontFacing,gl_ClipDistance,gl_FragColor,
    gl_FragData,gl_MaxDrawBuffers,gl_FragDepth,
    gl_PointCoord,gl_PrimitiveID,
    gl_MaxVertexAttribs,gl_MaxVertexUniformComponents,
    gl_MaxVaryingFloats,gl_MaxVaryingComponents,
    gl_MaxVertexOutputComponents,
    gl_MaxGeometryInputComponents,
    gl_MaxGeometryOutputComponents,
    gl_MaxFragmentInputComponents,
    gl_MaxVertexTextureImageUnits,
    gl_MaxCombinedTextureImageUnits,
    gl_MaxTextureImageUnits,
    gl_MaxFragmentUniformComponents,
    gl_MaxDrawBuffers,gl_MaxClipDistances,
    gl_MaxGeometryTextureImageUnits,
    gl_MaxGeometryOutputVertices,
    gl_MaxGeometryOutputVertices,
    gl_MaxGeometryTotalOutputComponents,
    gl_MaxGeometryUniformComponents,
  gl_MaxGeometryVaryingComponents,gl_DepthRange},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morecomment=[l][keywordstyle4]{\#},
}
\lstdefinestyle{GL}{
  tabsize=2,
  rulecolor=,
  basicstyle=\scriptsize,
  upquote=true,
  aboveskip={0.5\baselineskip},
  belowskip={1.5\baselineskip},
  columns=fixed,
  showstringspaces=false,
  extendedchars=true,
  breaklines=true,
  prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
  frame=single,
  showtabs=false,
  showspaces=false,
  showstringspaces=false,
  identifierstyle=\ttfamily,
  keywordstyle=\color[rgb]{1.0,0,0},
  keywordstyle=[1]\color[rgb]{0,0,0.75},
  keywordstyle=[2]\color[rgb]{0.5,0.0,0.0},
  keywordstyle=[3]\color[rgb]{0.127,0.427,0.514},
  keywordstyle=[4]\color[rgb]{0.4,0.4,0.4}}

\setcounter{biburllcpenalty}{7000}
\setcounter{biburlucpenalty}{8000}

%Information to be included in the title page:

\title{GPU Acceleration of the Material Point Method}

\author % (optional, for multiple authors)
{Fabian Meyer}

\institute % (optional)
{University of Koblenz}

\date{Kolloquium Computergrafik, 14 February 2019}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\section{Motivation on a GPU MPM Approach}

\subsection{A Brief MPM Overview: Do You Want to Build a Snowman?}

\begin{frame}

\frametitle{A Brief MPM Overview: Do You Want to Build a Snowman?}
A short historical summary of MPM:
\vfill
\begin{minipage}{0.43\paperwidth}
  \begin{itemize}
    \item Belongs to family of particle-in-cell(PIC) techniques \cite{evans1957particle}.
    \item Initial application to solids \cite{sulsky1995application} $\rightarrow$ MPM
    \item From research to production in \textit{Disney's} animation film \textit{Frozen} \cite{MPM:SNOW}.
    \item Avalanche research \cite{MPM:AVALANCHE}
  \end{itemize}
\end{minipage}
\begin{minipage}{0.35\paperwidth}
\includemedia[
  width=144,
  height=81,
  activate=pageopen,
  addresource=media/snowh264.mp4,
  flashvars={source=media/snowh264.mp4&autoPlay=true&loop=true&repeat=always}
  ]{\includegraphics{media/snow.jpg}}{VPlayer.swf}
  \tiny{Video result of my bachelor thesis on the simulation of snow \cite{Meyer2015}.}
\end{minipage}
\end{frame}

\begin{frame}
PIC ideas:
\vfill
\begin{minipage}{0.45\paperwidth}
\begin{itemize}
    \item Combine Lagrangian particles \& Eulerian grid    \item Particles store all information
\end{itemize}
\vfill
Typical PIC/MPM roundtrip:
\vfill
\begin{enumerate}
  \item Particle-to-grid(P2G) transfer to an unmoving grid
    \item Solve discretized governing equations on grid
    \item Grid-to-particle(G2P) transfer back to particles \& move them
    \item[$\Rightarrow$] \textbf{meshfree, non-empirical}
\end{enumerate}
\end{minipage}
\begin{minipage}{0.35\paperwidth}
	\includesvg[width=0.35\paperwidth]{media/pic.svg}
	\tiny{Transfers: Interpolation functions are defined over grid nodes.}
\end{minipage}
\end{frame}
\subsection{GPGPU for performance enthusiasts}
\begin{frame}
  \frametitle{GPGPU for performance enthusiasts}
  Why would('nt) you?
  \vfill
  \begin{columns}[T]
    \begin{column}{0.45\textwidth}
      \textbf{Drawbacks:}
    \begin{itemize}
      \item Interactivity much easier on CPU, but slow PCI-Bus communication
      \item Code is mostly written against GPU architecture
      \item A lot of strain on the programmer
    \end{itemize}
    \end{column}
    \begin{column}{0.45\textwidth}
      \textbf{Benefits:}
    \begin{itemize}
      \item Data is already on the GPU for rendering
      \item \textbf{Higher parallelization acceleration}
    \end{itemize}
    \end{column}
  \end{columns}
\end{frame}
\section{A Gentle Introduction to the MPM}

\subsection{Governing Equations: Conservation of Mass \& Momentum}

\begin{frame}
  \frametitle{Governing Equations: Conservation of Mass \& Momentum}
  \textbf{Conservation of mass}, continuum assumption holds.\\
  Lagrangian (moving with a particle ${}_0\boldsymbol{x}$):
  \begin{equation}
    {^t_0J}\rho(_0\boldsymbol{x},t) = {}\rho(_0\boldsymbol{x},0).
  \end{equation}
Eulerian (outside observer ${}_t\boldsymbol{x}$):
\begin{equation}
  \frac{\partial}{\partial t}\rho(_t\boldsymbol{x},t) = -  \vec{\nabla} \cdot (\rho(_t\boldsymbol{x},t) \boldsymbol{v}(_t\boldsymbol{x},t)).
\end{equation}
Lagrangian and Eulerian view measure differently but give same results. Equations are given in the strong form! \cite{MPM:COURSE}\cite{MIT:CONTINUUM_MECHANICS}
\end{frame}

\begin{frame}
  \textbf{Conservation of momentum}:\\
  Lagrangian (moving with a particle ${}_0\boldsymbol{x}$):
  \begin{equation}
  \rho(_0\boldsymbol{x},0)\boldsymbol{a}(_0\boldsymbol{x},t)
= \vec{\nabla} \cdot \boldsymbol{P}(_0\boldsymbol{x},t) + {\boldsymbol{f}} ^ {\text{body}}(_0\boldsymbol{x},t) {^t_0J}.
  \end{equation}
Eulerian (outside observer ${}_t\boldsymbol{x}$):
\begin{equation}
  \rho(_t\boldsymbol{x},t) \boldsymbol{a}(_t\boldsymbol{x},t) = \vec{\nabla} \cdot \boldsymbol{\sigma}(_t\boldsymbol{x},t) + \boldsymbol{f} ^ {\text{body}}(_t\boldsymbol{x},t)
\end{equation}
Solving this equation will tell us how the velocity fields $\boldsymbol{v}(_t\boldsymbol{x}),\boldsymbol{v}(_0\boldsymbol{x})$ change on the whole domain due to acceleration $\boldsymbol{a}$. This is important to advect particles accounting for all forces. \cite{MPM:COURSE}\cite{MIT:CONTINUUM_MECHANICS}
\end{frame}

\subsection{The Pretty Strong but Mathematically Weak Formulation}
\begin{frame}
\frametitle{The Pretty Strong but Mathematically Weak Formulation}

\textbf{Weak Formulation} (or Principle of Virtual Work):\\
Dot product equations with arbitrarily 'test functions' $\boldsymbol{q}$ and apply divergence theorem:
\begin{align}
\int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot  \Bigl[(_0\rho_0)(_0\boldsymbol{a}) - & {}_0\boldsymbol{f}^{\text{body}}{}_0^tJ\Bigr] d _0\boldsymbol{x}= \nonumber \\
										     &  \int _ { \partial \Omega ^ { t^n } } _t\boldsymbol{q} \cdot \boldsymbol{\sigma} d_t\boldsymbol{A} - \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}.
\end{align}
A strong solution is also a solution to the weak formulation. Leave out body forces(like gravity) and boundary condition (e.g. collisions) for now:
\begin{equation}
  \int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot (_0\rho_0)(_0\boldsymbol{a}) d _0\boldsymbol{x}= \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}.
\end{equation}
\end{frame}
\begin{frame}[fragile]
\textbf{Weak Derivative}:
\vfill
\begin{minipage}{0.49\textwidth}
  $y=|t|$ has weak derivative:
  \\\\
$ v = \left\{ \begin{array} { l l }
{ - 1 , } & { \text { if } t < 0 } \\
{ c , } & { \text { if } t = 0 } \\
{ 1 , } & { \text { if } t > 0 }
\end{array} \right.$
\\

\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{tikzpicture}
  \small
\datavisualization [school book axes,
		    y axis = {attribute = y, label},
		    visualize as line/.list={fx},
		    fx={style={color=blue}}
]

data [set=fx,format=function] {
      var x : interval [-1.5:1.5] samples 5;
      func y = abs(\value x);
      };
\end{tikzpicture}
\end{minipage}

\vfill
\begin{minipage}{0.49\textwidth}
  Heaviside step function has no weak derivate: \\\\
  $s= \left\{ \begin{array} { l l } { 0 , } & { \text { if } t < 0 } \\ { 1 , } & { \text { if } t \geq 0 } \end{array} \right.$
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{tikzpicture}
  \small
  \begin{axis}[axis lines = middle, width=5.1cm,height=3.5cm,xtick={-1,0,1},ytick={0,1},ymin=-0.3,ymax=1.25,xmin=-1.8,xmax=1.8,ylabel=$s$]
                \addplot[blue,mark=*,mark options={fill=white},samples at={-2.1,0}] {0};
                \addplot[blue,mark=*,samples at={0,2.1}] {1.0};
            \end{axis}
\end{tikzpicture}
\end{minipage}
\vfill
Allows for point loads, material discontinuities and more. \cite{bathe2006finite}
\end{frame}
\subsection{Discretization of Space and Time}
\begin{frame}
\frametitle{Discretization of Space and Time}
\textbf{Time discretization} with implicit midpoint scheme:
  \begin{equation}
    \frac { y ^ { n + 1 } - y ^ { n } } { \Delta t } = f ^ { n + \frac{1}{2} } = f \left( t ^ { n } + \frac{\Delta t}{2}, \frac{1}{2} y ^ { n } + \frac{1}{2} y ^ { n + 1 } \right)
  \end{equation}
\begin{itemize}
  \item \textbf{implicit} requires linear system solve $\Rightarrow$ more stable, larger time steps
  \item \textbf{midpoint} as it conserves gov. equations
\end{itemize}

  \begin{equation}
    \Rightarrow \int _ { \Omega ^ { t } } _t\boldsymbol{q} \cdot {}_t\rho(_t\boldsymbol{v}^{n+1}-{}_t\boldsymbol{v}^{n}) d _t\boldsymbol{x}= \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}^{n+\frac{1}{2}}  d_t\boldsymbol{x}.
  \end{equation}
\end{frame}

\begin{frame}[fragile]
\textbf{Space Discretization} is done in a Galerkin/FEM fashion with grid based interpolants $w_i$ with limited support. Here dyadic products
\begin{equation}
  w_i(\boldsymbol{x}) =w(\boldsymbol{x}-\boldsymbol{x}_i) = w(\frac{1}{h}(x-x_i))w(\frac{1}{h}(y-y_i)w(\frac{1}{h}(z-z_i))
\end{equation}
of cubic b-splines suffice:
\begin{equation}
w ( x ) = \left\{ \begin{array} { l l } { \frac { 1 } { 2 } | x | ^ { 3 } - | x | ^ { 2 } + \frac { 2 } { 3 } } & { 0 \leqslant | x | < 1 } \\ { \frac { 1 } { 6 } ( 2 - | x | ) ^ { 3 } } & { 1 \leqslant | x | < 2 } \\ { 0 } & { 2 \leqslant | x | } \end{array}. \right
\end{equation}
\begin{tikzpicture}[
  declare function={
    func(\x)= and(abs(\x) >= 0 , abs(\x)<1) * (pow(abs(\x),3)/2 - pow(abs(\x),2) +2/3)  +
	      and(abs(\x) >= 1 , abs(\x)<2) * (1/6* pow((2-abs(\x)),3))
   ;
  }
]
\begin{axis}[
  width=10cm,height=4cm,
  axis x line=middle, axis y line=middle,
  ymin=-0.1, ymax=1, ytick={0,1}, ylabel=$w$,
  xmin=-2.5, xmax=2.5, xtick={-2,...,2}, xlabel=$x$,
  domain=-2.5:2.5,samples=101, % added
]

\addplot [blue,thick] {func(x)};
\end{axis}
\end{tikzpicture}
\end{frame}
\begin{frame}
Thus set in:
$$
{}_t\boldsymbol{q}(\boldsymbol{x},t^n) = \sum_i \boldsymbol{e}_iw_i(\boldsymbol{x}),
{}_t\boldsymbol{v}^{n(+1)}(\boldsymbol{x}) = \sum_j \boldsymbol{v}^{n(+1)}_jw_j(\boldsymbol{x}).
$$
Combine it with numerical integration where the particles function as quadrature points \cite{Steffen}:

\begin{minipage}{0.6\textwidth}

$$
g _ { i } = \int _ { \Omega } g ( \boldsymbol{x} ) w _ { i } ( \boldsymbol{x} ) d \boldsymbol{x}
$$
\begin{equation}
  \approx \sum _ { p } g _ { p }  w_i(\boldsymbol{x}_p) V _ { p }.
\end{equation}
due to integration by midpoint rule:
\begin{equation}
\int _ { \Omega } f ( x ) d x \approx \sum _ { i = 1 } ^ { N } f \left( x _ { i } \right) h _ { i }.
\end{equation}

\end{minipage}
\begin{minipage}{0.3\textwidth}
  \begin{tikzpicture}[/pgf/declare function={f=-.1*(x)*(x-8);}]
\pgfplotsset{
    integral segments/.code={\pgfmathsetmacro\integralsegments{#1}},
    integral segments=3,
    integral/.style args={#1:#2}{
        ybar interval,
        domain=#1+((#2-#1)/\integralsegments)/2:#2+((#2-#1)/\integralsegments)/2,
        samples=\integralsegments+1,
        x filter/.code=\pgfmathparse{\pgfmathresult-((#2-#1)/\integralsegments)/2}
    }
}
\begin{axis}[
    ylabel=$f$,
    width =5cm,height=5cm,
    domain=0:6,
    samples=100,
    axis lines=middle
]
\addplot [thick] {f};
\addplot [
    red,
    integral=0:6
] {f};
\end{axis}
\end{tikzpicture}
\end{minipage}
\end{frame}
  \subsection{Velocity Fields: APIC-Transfers}
\begin{frame}
  \frametitle{Velocity Fields: APIC-Transfers}
  \begin{minipage}{0.64\textwidth}
  PIC-transfer
  \begin{enumerate}
    \item $(m\boldsymbol{v})_i^n = \sum_p w_{ip}^nm_p\boldsymbol{v}^n_p$
    \item $\boldsymbol{v}_i^n = \frac{(m\boldsymbol{v})_i^n}{m_i^n}$
    \item $\boldsymbol{v}_{p,{PIC}}^{n+1} = \sum_i w^n_{ip}\boldsymbol{v}_i^{n+1}$
  \end{enumerate}
\vspace*{15}
APIC-transfers add a local velocity field $\boldsymbol{C}^n_p$ around $\boldsymbol{v}^n_p$:
\\\vspace*{-10}\\
$ ( m \boldsymbol{v} ) _ { i } ^ { n } = \sum _ { p } w _ { i p } ^ { n } m _ { p } \left( \boldsymbol{v} _ { p } ^ { n } + \boldsymbol{C} _ { p } ^ { n } ( \boldsymbol{x} _ { i } ^ { n } - \boldsymbol{x} _ { p } ^ { n }))$
\end{minipage}
\begin{minipage}{0.30\textwidth}
\begin{minipage}{1.0\textwidth}
  \small
  \includesvg[width=1.0\textwidth]{media/pic.svg}
  P2G-transfer $\sum_p$
\\
\end{minipage}
\begin{minipage}{1.0\textwidth}
  \small
  \includesvg[width=1.0\textwidth]{media/g2p.svg}
  G2P-transfer $\sum_i$
\end{minipage}
\end{minipage}
\end{frame}
\section{A MPM Guide on GPGPU Pitfalls and Optimizations}
\subsection{Layout of the data: SoA vs. AoS}
\begin{frame}[fragile]
\frametitle{Layout of the data: SoA vs. AoS}

\begin{minipage}{0.31\textwidth}
\begin{minipage}{0.95\textwidth}
Coalescing:
\vspace*{23}
\end{minipage}
\tiny
\begin{minipage}{0.95\textwidth}
  SoA-Layout
\begin{lstlisting}[language={GLSL},style={GL},basicstyle=\tiny]
struct Container{
  vec4 positions[n];
  vec4 velocities[n];
} Particles;
\end{lstlisting}
\vspace*{29}
\end{minipage}
\begin{minipage}{0.95\textwidth}
  AoS-Layout
\begin{lstlisting}[language={GLSL},style={GL},basicstyle=\tiny]
struct Particle {
  vec4 position;
  vec4 velocity;
} Particles[n];
\end{lstlisting}
\end{minipage}
\end{minipage}
\begin{minipage}{0.64\textwidth}
  \tiny
  \includesvg[svgpath =../images/,width=\textwidth]{soa}
  Structures of arrays(SoA)
\\
  \includesvg[svgpath =../images/,width=\textwidth]{aos}
  Arrays of structures(AoS)

\end{minipage}
\end{frame}
\begin{frame}
Nvidia Nsight\cite{NVIDIA:METRICS} now offers metrics to identify bottlenecks:
\vspace{-5}
\begin{table}
  \footnotesize
  \begin{tabular}{ | l | l |}    \hline
    \textbf{Metric}   & \textbf{Description} \\ \hline
    VRAM SOL\%        & memory througput w.r.t. to hardware limit\\\hline
    SM SOL\%          & instruction throughput\\\hline
    L2 SOL\%          & L2-cache throughput\\\hline
    Tex SOL\%         & L1-cache throughput\\\hline
    SM Issue Util.\%  & amount of cycles an instr. was issued\\\hline
  \end{tabular}
\end{table}
\\
\vspace{5}
A simple map(\texttt{y=length(x)}) shader on 1024$\times$1024 Elements SoA vs. AoS differences:
\vspace{-15}
\begin{table}
  \footnotesize
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Layout &  $\Delta t_c (\mu s)$	& Speedup &VRAM     & SM     & L2     & SM Issue Util. \\\hline
    AoS(1 instr.)         & 243                   & -       &77.7\%   & 7.3\%  & 30.3\% & 6.8\% \\\hline
    SoA(1 instr.)         & \textbf{120}                   & 2.26x   &75.4\%   & \textbf{14.3}\% & 29.4\% & \textbf{14.0\%} \\\hline
    AoS(2 instr.)         & 275                   & -        &61.3\%           & \textbf{41.8}\% & \textbf{53.8}\% & 48.9\% \\\hline
    SoA(2 instr.)         & 240                   & 1.16x    &75.4\%  & 29.4\%          & 20.0\% & 62.3\% \\
    \hline
\end{tabular}
\end{table}
$\Rightarrow$ SoA increases coalescing for non-random access.
\end{frame}
\subsection{Parallel Reduction \& Scan}
\begin{frame}
\frametitle{Parallel Reduction \& Scan}
Assuming an associative \texttt{binary\_op(x,y)}:= $x \circ y$ , a neutral element $e$ of the \texttt{binary\_op}, and an array of values $\left[a_0,a_1, ... ,a_n\right]$.\\\vspace{10}
\begin{itemize}
  \item \textbf{Parallel reduction} computes the value:
\begin{equation}
  r = a_0 \circ a_1 \circ ... \circ a_n.
\end{equation}
\item (Exclusive) \textbf{scan} computes the array:
\begin{equation}
  \left[e , a_0 , \left(a_0 \circ a_1\right), \left(a_0 \circ a_1 \circ a_2\right), \ ... \ ,\left(a_0\circ a_1 \circ a_2 \circ ...\circ a_{n-1}\right)\right].
\end{equation}
\end{itemize}
\\\vspace{5}
Here, only shared memory approaches without (NVIDIA exclusive) warp shuffle operations.
\end{frame}
\begin{frame}
\textbf{Shared Memory Bank Conflicts}:
\\\vspace{10}
\begin{minipage}{0.94\textwidth}
\begin{minipage}{0.5\textwidth}
  \tiny
  \includesvg[svgpath =../images/,width=1.0\textwidth]{sequential}
  Sequential Addressing
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \tiny
  \includesvg[svgpath =../images/,width=1.0\textwidth]{interleaved}
  Interleaved Addressing
\end{minipage}
\end{minipage}
\\\vspace{10}
\\
Interleaved addressing causes bank conflicts (Short Scoreboard activity) $\Rightarrow$ padding needed.
\vspace{-12}
\begin{table}
    \footnotesize
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method         &  $\Delta t_c$ & Speedup &VRAM              & SM              & Sel. Warp-Stall Reas.\\\hline
    Interl. no padd.    & 305                   & -       &23.0\%          & \textbf{60.9}\%  & S. Scoreb.(17.2\%)\\\hline
    Sequential     & 141                   & 2.16x   &\textbf{49.8}\%   & 37.1\%         & \textbf{S. Scoreb.(2.0\%)} \\\hline
  \end{tabular}
    \caption{\footnotesize Parallel reduction on $1024\times1024$ vectors with \texttt{y=length(x)} as input.}
\end{table}
\vspace{-10}
More elements than thread group size require pyramid schemes.
\end{frame}

\begin{frame}

\textbf{Sequential work}: multiple elements per thread.\begin{itemize}
  \item Memory latency hiding (Long Scoreboard up)
  \item Higher reduction factor each dispatch $\Rightarrow$ Less global memory indirections
  \item Unrolling loops can help but adds register pressure.
\end{itemize}
\begin{table}
  \footnotesize
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method         &  $\Delta t_c$ & Speedup &VRAM              & SM              & Sel. Warp-Stall Reas.\\\hline
    Sequential     & 141                   & 2.16x   &49.8\%   & 37.1\%         & S. Scoreb.(2.0\%) \\\hline\hline
    Seq. (2x)      & 100                   & 3.05x   &69.5\%            & 26.2\%         & L. Scoreb.(80.1\%)\\\hline
    Seq. (128x)    & 98                    & 3.1x    &\textbf{72.9}\%   & 16.9\%         & \textbf{L. Scoreb.(84.4\%)}\\\hline
    Seq. (256x)    & 101                   & 3.0x    &66.4\%            & 14.6\%         & L. Scoreb.(76.9\%)\\
    \hline
  \end{tabular}
  \caption{\footnotesize Parallel reduction on $1024\times1024$ vectors with \texttt{y=length(x)} as input. Methods have 504, 8, 4 thread groups, respectively. A GTX970 has 13 SMs.}
\end{table}
\vspace{-10}
Scan is similar but cannot profit as much from sequential work having to keep multiple elements in register memory.
\end{frame}
\subsection{Binning \& Counting Sort: Where Are You?}
\begin{frame}
\frametitle{Binning \& Counting Sort: Where Are You?}

Grid node does not know its neighboring particles $\Rightarrow$ Binning.
\\\vspace{10}
Binning combines nicely with \textbf{Counting Sort:}\\
\begin{minipage}{0.55\textwidth}
\begin{enumerate}
  \item \textbf{Binning}: Per node counting.
  \item \textbf{Scan}: Computes new memory offset for particles.
  \item \textbf{Reordering}: Give back indexing list or do deep copy.
\end{enumerate}
\end{minipage}
\begin{minipage}{0.40\textwidth}
  \footnotesize
 \includesvg[svgpath =../images/,width=\textwidth]{binning}
 \caption{\footnotesize Binning: Particle adds one
to counter after storing it as offset.}
 \end{minipage}
\end{frame}

\begin{frame}
Sorting can dramatically increase workload performance of subsequent steps for neighboring queries:
\begin{enumerate}
  \item Deep sorted accesses are now \textbf{coalesced}. \label{it:sort_1}
  \item \textbf{Data reuse} due to L2-Cache and/or shared memory. \label{it:sort_2}
\end{enumerate}
\textbf{Double buffer particles} to use last sorted state as input for new sorting to profit from item 1 and 2!
\begin{table}
  \footnotesize
  \begin{tabular}{ | l | r | r | r | r| r |r|}    \hline
    Ordering            &  $\Delta t_c (\mu s)$ & Speedup &VRAM              & SM      & L2     & L2-Hit \\\hline
    Random              &  1,516                & -       &25.0\%            & 3.4\%   & 9.1\%  & 10.8\% \\\hline
    Deep sorted           &    218                & 6.95x   &\textbf{75.3\%}   & 24.4\%  & \textbf{35.0\%} & \textbf{37.8\%} \\
    \hline
 \end{tabular}
\caption{\footnotesize Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid.}\label{table:bin}
\end{table}
\end{frame}
\subsection{The MPM Specific Transfers}
\begin{frame}
\frametitle{The MPM Specific Transfers}
All MPM operations belong to one of those parallelization schemes:
\begin{itemize}
 \item 1 thread : 1 particle: $\square_p = \square_p \circ \square_p \circ ... \circ \square_p$.
 \item 1 thread : 1 node: $\square_p = \square_p \circ \square_p \circ ... \circ \square_p$.
 \item G2P-transfer: $\square_p = \sum_i \square_i \circ \square_{ip}.$
 \item P2G-transfer: $\square_i = \sum_p \square_p \circ \square_{ip}.$
\end{itemize}
\vfill
MPM-Transfers are executed \textbf{multiple times per physical frame} with varying numbers of variables and mathematical operations.\\\vspace{5}
$\Rightarrow$ Preprocessing steps only need to be done \textbf{once per physical frame}. Sorting already introduced as one of these.
\end{frame}
\begin{frame}
\begin{minipage}{\textwidth}
  G2P-Transfer: $\square_p = \sum_i \square_i \circ \square_{ip}.$\\
  \includesvg[svgpath =../images/,width=\textwidth]{block_transfer_inv_par}
  Pull-Operations read!
\end{minipage}
\end{frame}
\begin{frame}
Similar to filter or stencil operations on the GPU:
\begin{itemize}
  \item 1 node : 1 thread, split grid into blocks corr. to thread groups.
  \item Interpolation function however dependent on particle position.
  \item Needs to be rerun for every particle in the cell.
\end{itemize}
\vfill
Typical setup of transfers:
\begin{enumerate}
  \item Initialize shared memory (pull: with nodes from global memory).
  \item Perform transfers.
  \item Write back to global memory (push: global atomics since writes on halo).
\end{enumerate}
\end{frame}
\begin{frame}
\begin{minipage}{\textwidth}
  P2G-transfer: $\square_i = \sum_p \square_p \circ \square_{ip}.$\\
  \includesvg[svgpath =../images/,width=\textwidth]{block_transfer_par}
  Pull-Operations read, Push-operations write!\\
  $\Rightarrow$ Race conditions.
\end{minipage}
\end{frame}
\begin{frame}
  Typically simulation domain(grid) much bigger than simulation model. Filtering inactive blocks as a preprocess improves performance.
\begin{itemize}
  \item Block is active if any cell counter is active.
  \item Cell counter is active if it has at least one particle.
\end{itemize}
\vfill
\begin{minipage}{0.9\textwidth}
  \footnotesize
  \includesvg[svgpath =../images/,width=\textwidth]{indirect_dispatch}
\end{minipage}
\end{frame}
\begin{frame}
\begin{itemize}
  \item Block and its halo are always target of shared memory operations $\Rightarrow$ P2G-Pull low occupancy.
  \item Batching multiple particles can increase performance due to hiding synchronization, unroll!
  \item Transfers respect shared memory bank conflicts fully.
  \item Warp divergence for varying cell counts.
\end{itemize}

{\footnotesize
\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM     \\\hline
    global                 &   44,442             &-               &4.6 \%&34.4\%&7.7\%  \\\hline
    global sorted          &   20,484		  &2.21x           &7.0 \%&\textbf{44.0}\%&16.1\% \\\hline
    P2G-sync              &    \textbf{2,595}    &17.47x          &5.9\%&7.6\%  &\textbf{67.0}\%\\\hline
 \end{tabular}
\end{table}
P2G-transfers of one uniformly million particles with 4 particles per cell with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. Block size is (8,4,4).
}

\end{frame}
\section{Delving Deeper: Further Opportunities}
\subsection{A comparison to \cite{Gao:2018}}
\begin{frame}
  \frametitle{A comparison to \cite{Gao:2018}}

Largely same decision making:\\
\begin{table}
  \begin{tabular}{|l|c|c|}
    \hline
           & Me              & \cite{Gao:2018} \\\hline
    Sort   & Count/Histogram & Count/Histogram \\
	   & for each var.   & sel. variables  \\\hline
    Filtering domain& Filter-op.& Sparse Grid structure \\\hline
    Transfers        & Shared mem. only & Warp-shuffle op.\\\hline
  \end{tabular}
\end{table}
Warp shuffle allows for fast parallel segmented reduction of cells of varying counts.\\
$\Rightarrow$ Solves warp divergence and shared memory issues mostly, thread groups now correspond to particles. Faster in these instances but NVIDIA only.

\end{frame}

\selectlanguage{english}
\printbibliography
\end{document}